Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures

### 2012-01-04 ltx/paper.tex
more work on SMC inference algorithm

\documentclass[smallcondensed, final]{svjour3}
% \documentclass[smallcondensed, final]{svjour3}[1/1/2012]

@@

@@

\bold{x}_{i} | c_{i}, \theta_{1:K_{t}, t} &\sim F(\theta_{c_{i}, t})

@@

        \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn Dependent Dirichlet Process Mixture. Note that all observations at time $t$, $\bold{x}_{1:N,t}$, and their associated assignments $c_{1:N,t}$ are denoted respectively as $x_{t}$ and $c_{t}$ (this also holds for time step $t-1$).}

@@

At a given time $t$, we model each $\bold{x} \in \bold{X}$ as a draw from the product of a multivariate normal and multinomial distribution
\begin{equation}
P(\bold{x}|\theta) = \mathcal{N}(\bold{x}^{s} | \boldsymbol{\mu}, \Sigma)  \mathcal{M}n(\bold{x}^{c} | \bold{p})
\end{equation}
where $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ denote the parameters of a cluster at time $t$, with mean $\boldsymbol{\mu} \in \mathbb{R}^{2}$, covariance matrix $\Sigma \in \mathbb{R}^{2} \times \mathbb{R}^{2}$, and discrete probability vector $\bold{p} = (p_{1}, \ldots, p_{V})$ such that $\sum_{i=1}^{V}p_{i} = 1$. Also, $\mathcal{N}$ denotes the multivariate normal distribution and $\mathcal{M}n$ denotes the multinomial distribution,

@@

$\mathbb{G}_{0}$ denotes the base distribution of the time-dependent Dirichlet process mixture; it also serves as a prior distribution for the parameters $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ present in the likelihood. We make use of conjugate priors in the base distribution to allow for more efficient computation. Specifically, in the experiments carried out in this paper, a normal-inverse-Wishart prior is placed on the multivariate normal parameters $\{ \boldsymbol{\mu}, \Sigma \}$ , and a Dirichlet prior is placed on the multinomial parameter $ \{  \bold{p}  \} $ (where the normal-inverse-Wishart is a conjugate prior for the multivariate normal component of the likelihood and the Dirichlet is a conjugate prior for the multinomial component). The prior can thus be written
\begin{equation} \label{basedistro}
\mathbb{G}_{0}(\theta) = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
\end{equation}
where $\mathcal{N}i\mathcal{W}$ denotes the normal-inverse-Wishart distribution, $\mathcal{D}ir$ denotes the Dirichlet distribution, and the prior has the hyperparameters $\boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0}$ and $\bold{q}_{0}$.  \willie{reference text with good background on NiW and Dir distributions}

@@

As described in Section~\ref{sec:gpuddpm}, $\mathbb{G}_{0}$ must be the invariant distribution of $P(\theta_{t} | \theta_{t-1})$ in order for the the cluster parameters to remain marginally distributed according to the base distribution. In other words, the transition kernel must satisfy
\begin{equation}
\int \mathbb{G}_{0}(\theta_{t-1})P(\theta_{t} | \theta_{t-1}) d\theta_{t-1} = \mathbb{G}_{0}(\theta_{t})
\end{equation}
for a given cluster $\theta$. One way to achieve this is through the use of auxiliary variables. Auxiliary variables are a set of $M$ variables $\bold{z}_{t} = (z_{t,1}, \ldots, z_{t,M})$ associated with a each cluster $\theta$ at time $t$ that satisfy
\begin{eqnarray}
% P(\theta_{k,t} | \theta_{k,t-1}) = \int P(\theta_{k,t} | \bold{z}_{k,t}) P(\bold{z}_{k,t} | \theta_{k,t-1}) d \bold{z}_{k,t}
P(\theta_{t} | \theta_{t-1}) = \int P(\theta_{t} | \bold{z}_{t}) P(\bold{z}_{t} | \theta_{t-1}) d \bold{z}_{t}  %\\
% P(\theta', \bold{z}_{t}) = p(\bold{z}_{t}|\theta') \mathbb{G}_{0}(\theta')
\end{eqnarray}

@@

For each cluster, we introduce $M$ auxiliary variables $z_{t, 1}, \ldots, z_{t, M}$ at time $t$ that are each drawn from the product of a multivariate normal and multinomial when conditioned on the associated cluster parameters $\theta_{t} = \{ \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} \}$:
\begin{equation} \label{transker1}
z_{t, m} | \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t}  \sim  \mathcal{N}(\boldsymbol{\mu}_{t}, \Sigma_{t}) \mathcal{M}n(\bold{p}_{t})   \hspace{15pt}   
\forall m \in \{ 1, \ldots, M \}
\end{equation}
To satisfy the above criteria for auxiliary variables, at each time $t$ we specify the dependencies of a given cluster on its associated set of auxiliary variables by
\begin{equation} \label{transker2}
\boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} | \bold{z}_t  \sim  \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{M}, \kappa_{M}, \nu_{M}, \Lambda_{M})  \mathcal{D}ir(\bold{q}_{M})
\end{equation}
where $\boldsymbol{\mu}_{M}, \kappa_{M}, \nu_{M}, \Lambda_{M},$ and $\bold{q}_{M}$ are parameters given by
\willie{these are given in next section, need to consider if I actually want to repeat them here}

@@

The multivariate normal-multinomial GPUDDPM implemented in this study has a number of parameters, which can be tuned to increase the efficacy of object detection and tracking. Object apperance parameters include\\\\
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$\boldsymbol{\mu}_{0}$  &  The mean prior. In experiments performed in Section~\ref{sec:experiments} the data was recentered to the origin, and this parameter was set to $(0,0)$ \\
$\kappa_{0}$  &  scale factor of the variance of the prior on the mean\\
$\nu_{0}$  &  scale factor of the variance of the prior on the covariance\\
$\Lambda_{0}$  &  shape factor of the prior on the covariance\\
$q_{0}$  &  scale factor of the prior on the multinomial counts
\end{tabular}
\end{center} \vspace{5mm}
The following parameter dictates characteristics involving the movement of objects
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$M$  &  Number of auxiliary variables. A large number will produce lower variation in position (better for slower objects) and a low number will produce higher variation in position.
\end{tabular}
\end{center} \vspace{5mm}
Additionally, one can tune the model's tendency to detect new objects and maintain the existence of these objects (or the rate at which objects become inactive or leave the video) with the following parameters
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$\alpha$  &  The granularity parameter for the Dirichlet Process. A higher value will increase the tendency for new objects to be detected.\\
$\rho$  &  The deletion parameter. A higher value will give objects an increased tendency to die off.
\end{tabular}
\end{center}
\willie{Discuss choosing parameters}

@@

The generative model associated with the GPUDDPM is illustrated in Section~\ref{sec:gpuddpm}. This representation includes cluster size $m$ (the number of observations assigned to a given cluster), which is standard in DPM constructions. To capture time-dependence, the GPUDDPM allows clusters to `lose assigned observations' in a deletion step, which modifies the sizes $m$ over time. Instead of incorporating cluster size variables $m$ directly, we can formulate an equivalent model which makes use of deletion variables $d$, which represent the time when observations are lost from a cluster . At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables. In particular, we introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed. With these variables, we can now express the $m_{k}$ value at time $t$ as
\begin{equation}
\label{compute_clust_size}
m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
\end{equation}
where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise. Additionally, for an observation $x_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
\begin{equation}
\label{del_rho_form}
l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
\end{equation}
The MCMC sampler described in this section operates in this alternate representation, which we can formulate, for each time step $t \in \{1, \ldots, T\}$ and observation $i \in \{ 1, \ldots, N_{t} \}$ at time $t$, as
\begin{align}
\begin{split}
	 \bold{x}_{i} | c_{i}, \theta_{c_{i},t} &\sim \mathcal{N}(\boldsymbol{\mu}_{c_{i},t}, \Sigma_{c_{i},t})  \mathcal{M}n(\bold{p}_{c_{i},t}) \\
	c_{i} | c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha  &\sim  \text{CRP}(c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha) \\
	d_{i} | \rho  &\sim \text{Geo}(\rho) + i + 1\\
	\theta_{k, t} | \theta_{k, t-1}, \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0}  &\sim 
\begin{cases}
	\text{TK} \hspace{12pt} \text{if } k \in \{ 1, \ldots, K_{t-1}  \} \\
	\mathbb{G}_{0}(\boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0})  \hspace{15pt} \text{if } k = K_{t-1} + 1
\end{cases}
\end{split}
\end{align}
where TK stands for the transition kernel, given in \eqref{transker1} and \eqref{transker2}, and the base distribution $\mathbb{G}_{0}$ is given in \eqref{basedistro}. This formulation of the GPUDDPM is also used in \cite{gasthaus_thesis} and \cite{caron_2007}. The associated graphical model for this formulation is given in Figure~\ref{fig:gpuddpm_gm_2}.
\begin{figure}[h]
        \center{\includegraphics[width=100mm]{../img/gpuddp_gm_3.pdf}}
        \caption{Graphical model of the deletion variable and auxiliary variable transition kernel representation of the GPUDDPM}
        \label{fig:gpuddpm_gm_2}
\end{figure}

@@

\subsubsection{Sampling Cluster Parameters $\theta_{k, t}$}
The conjugacy present in chosen distributions and auxiliary variables in the transition kernel allow for posterior samples of the cluster parameters to be easily gained, and gives us
% \begin{eqnarray}
% P(\theta_{k,t} | state) && = P(\bold{x}_{i} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
%  && = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
% \end{eqnarray}
\begin{align}
\begin{split}
P(\theta_{k,t} | state) & = P(\bold{x}_{i} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
 & = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{k,t}, \Sigma_{k,t} | \boldsymbol{\mu}_{N}, k_{N}, v_{N}, \Lambda_{N})  \mathcal{D}ir( \bold{p}_{k,t} | \bold{q}_{N})
\end{split}
\end{align}
Where the parameters in the above distribution are given when the data at time $t$, auxiliary variables $z_{k,t-1,1:M}$ at time $t-1$, and auxiliary variables $z_{k,t,1:M}$ at time $t$, are taken to be the observations in the following Bayesian updates
\begin{eqnarray}
k_{N} &=& k_{0} + N \\
v_{N} &=& v_{0} + N \\
\boldsymbol{\mu}_{N} &=& \frac{k_{0}}{k_{0}+N} \boldsymbol{\mu}_{0}  +  \frac{N}{k_{0}+N} \overline{\bold{x}}^{s}\\
\Lambda_{N} &=& \Lambda_{0} + S_{\bold{x}^{s}}\\
\bold{q}_{N} &=& \bold{q}_{0} + \sum_{i=1}^{N} \bold{x}^{c}
\end{eqnarray}
where $N$ is the number of observations, $\{ \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{N} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{x}^{s}$ and $\bold{x}^{c}$ respectively denote the spatial and color features of the observations, and $\overline{\bold{x}}$ and $S_{\bold{x}}$ respectively denote the sample mean and sample covariance for a set of observations $\bold{x}$, defined to be
\begin{eqnarray}
\overline{\bold{x}}  &=&  \left( \sum_{i=1}^{N} \bold{x}_{i} \right) / N\\
S_{\bold{x}}  &=&  \sum_{i=1}^{N} (\bold{x}_{i} - \overline{\bold{x}}) (\bold{x}_{i} - \overline{\bold{x}})^{T}
\end{eqnarray}

@@

This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference, which provides another method to carry out unsupervised detection and tracking. The SMC inference algorithm is given in Algorithm~\ref{alg:SMC}. The algorithm operates by

@@

>>>>

\documentclass[smallcondensed, draft]{svjour3}
% \documentclass[smallcondensed, final]{svjour3}[1/1/2012]

@@

\label{CRP}

@@

\bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t} &\sim F(\theta_{c_{i,t}, t})

@@

        \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn Dependent Dirichlet Process Mixture. Note that all observations at time $t$, $\bold{x}_{1:N,t}$, and their associated assignments $c_{1:N,t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (this also holds for time step $t-1$).}

@@

At a given time $t$, we model each $\bold{x} \in \bold{X}$ as a draw from the product of a multivariate normal and multinomial distribution
\begin{equation}
\label{likelihood}
P(\bold{x}|\theta) = \mathcal{N}(\bold{x}^{s} | \boldsymbol{\mu}, \Sigma)  \mathcal{M}n(\bold{x}^{c} | \bold{p})
\end{equation}
where $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ denote the parameters of a cluster at time $t$, with mean $\boldsymbol{\mu} \in \mathbb{R}^{2}$, covariance matrix $\Sigma \in \mathbb{R}^{2} \times \mathbb{R}^{2}$, and discrete probability vector $\bold{p} = (p_{1}, \ldots, p_{V})$ such that $\sum_{i=1}^{V}p_{i} = 1$. Also, $\mathcal{N}$ denotes the multivariate normal distribution and $\mathcal{M}n$ denotes the multinomial distribution,

@@

$\mathbb{G}_{0}$ denotes the base distribution of the time-dependent Dirichlet process mixture; it also serves as a prior distribution for the parameters $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ present in the likelihood. We make use of conjugate priors in the base distribution to allow for more efficient computation. Specifically, in the experiments carried out in this paper, a normal-inverse-Wishart prior is placed on the multivariate normal parameters $\{ \boldsymbol{\mu}, \Sigma \}$ , and a Dirichlet prior is placed on the multinomial parameter $ \{  \bold{p}  \} $ (where the normal-inverse-Wishart is a conjugate prior for the multivariate normal component of the likelihood and the Dirichlet is a conjugate prior for the multinomial component). The prior can therefore be written
\begin{equation}
\label{basedistro}
\mathbb{G}_{0}(\theta) = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
\end{equation}
where $\mathcal{N}i\mathcal{W}$ denotes the normal-inverse-Wishart distribution, $\mathcal{D}ir$ denotes the Dirichlet distribution, and the prior has the hyperparameters $\boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0}$ and $\bold{q}_{0}$.  \willie{reference text with good background on NiW and Dir distributions}

@@

As described in Section~\ref{sec:gpuddpm}, $\mathbb{G}_{0}$ must be the invariant distribution of $P(\theta_{t} | \theta_{t-1})$ in order for the the cluster parameters to remain marginally distributed according to the base distribution. In other words, the transition kernel must satisfy
\begin{equation}
\int \mathbb{G}_{0}(\theta_{t-1})P(\theta_{t} | \theta_{t-1}) d\theta_{t-1} = \mathbb{G}_{0}(\theta_{t})
\end{equation}
for a given cluster with parameters $\theta$. One way to achieve this is through the use of auxiliary variables. Auxiliary variables are a set of $M$ variables $\bold{z}_{t} = (z_{t,1}, \ldots, z_{t,M})$ associated with a each cluster at time $t$ that satisfy
\begin{eqnarray}
% P(\theta_{k,t} | \theta_{k,t-1}) = \int P(\theta_{k,t} | \bold{z}_{k,t}) P(\bold{z}_{k,t} | \theta_{k,t-1}) d \bold{z}_{k,t}
P(\theta_{t} | \theta_{t-1}) = \int P(\theta_{t} | \bold{z}_{t}) P(\bold{z}_{t} | \theta_{t-1}) d \bold{z}_{t}  %\\
% P(\theta', \bold{z}_{t}) = p(\bold{z}_{t}|\theta') \mathbb{G}_{0}(\theta')
\end{eqnarray}

@@

For each cluster, we introduce $M$ auxiliary variables $z_{t, 1}, \ldots, z_{t, M}$ at time $t$ that are each drawn from the product of a multivariate normal and multinomial when conditioned on the associated cluster parameters $\theta_{t} = \{ \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} \}$:
\begin{equation}
\label{transker1}
z_{t, m} | \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t}  \sim  \mathcal{N}(\boldsymbol{\mu}_{t}, \Sigma_{t}) \mathcal{M}n(\bold{p}_{t})   \hspace{15pt}   
\forall m \in \{ 1, \ldots, M \}
\end{equation}
To satisfy the above criteria for auxiliary variables, at each time $t$ we specify the dependencies of a given cluster on its associated set of auxiliary variables by
\begin{equation}
\label{transker2}
\boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} | \bold{z}_t  \sim  \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{M}, \kappa_{M}, \nu_{M}, \Lambda_{M})  \mathcal{D}ir(\bold{q}_{M})
\end{equation}
where $\boldsymbol{\mu}_{M}, \kappa_{M}, \nu_{M}, \Lambda_{M},$ and $\bold{q}_{M}$ are parameters given by
\begin{eqnarray}
\kappa_{M} &=& \kappa_{0} + M \\
\nu_{M} &=& \nu_{0} + M \\
\boldsymbol{\mu}_{M} &=& \frac{\kappa_{0}}{\kappa_{0}+M} \boldsymbol{\mu}_{0}  +  \frac{M}{\kappa_{0}+M} \overline{\bold{z}_t}^{s}\\
\Lambda_{M} &=& \Lambda_{0} + S_{\bold{z}_{t}^{s}}\\
\bold{q}_{M} &=& \bold{q}_{0} + \sum_{m=1}^{M} z_{t,m}^{c}
\end{eqnarray}
where $M$ is the number of auxiliary variables, $\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{z}^{s}$ and $\bold{z}^{c}$ respectively denote the spatial and color features of an auxiliary variable $\bold{z}$, and $\overline{\bold{z}}$ and $S_{\bold{z}}$ respectively denote the sample mean and sample covariance for a set $\bold{z} = \{ z_{1}, \ldots, z_{M} \}$ (of auxiliary variables, in this case), which we can define as
\begin{eqnarray}
\label{samplemean}
\overline{\bold{z}}  &=&  \left( \sum_{m=1}^{M} z_{m} \right) / M\\
\label{samplecov}
S_{\bold{z}}  &=&  \sum_{m=1}^{M} (z_{m} - \overline{\bold{z}}) (z_{m} - \overline{\bold{z}})^{T}
\end{eqnarray}

@@

The multivariate normal-multinomial GPUDDPM implemented in this study has a number of parameters, which can be tuned to increase the efficacy of object detection and tracking. Object apperance parameters include\\\\
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$\boldsymbol{\mu}_{0}$  &  The mean prior. In experiments performed in Section~\ref{sec:experiments} the data was recentered to the origin, and this parameter was set to $(0,0)$ \\
$\kappa_{0}$  &  scale factor of the variance of the prior on the mean\\
$\nu_{0}$  &  scale factor of the variance of the prior on the covariance\\
$\Lambda_{0}$  &  shape factor of the prior on the covariance\\
$\bold{q}_{0}$  &  scale factor of the prior on the multinomial counts
\end{tabular}
\end{center} \vspace{5mm}
The following parameter dictates characteristics involving the movement of objects
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$M$  &  Number of auxiliary variables. A large number will produce lower variation in position (better for slower objects) and a low number will produce higher variation in position.
\end{tabular}
\end{center} \vspace{5mm}
Additionally, one can tune the model's tendency to detect new objects and maintain the existence of these objects (or the rate at which objects become inactive or leave the video) with the following parameters
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$\alpha$  &  The granularity parameter for the Dirichlet Process. A higher value will increase the tendency for new objects to be detected.\\
$\rho$  &  The deletion parameter. A higher value will give objects an increased tendency to die off.
\end{tabular}
\end{center}
\willie{Discuss choosing parameters}

@@

The generative model associated with the GPUDDPM is illustrated in Section~\ref{sec:gpuddpm}. This representation includes cluster size $m$ (the number of observations assigned to a given cluster), which is standard in DPM constructions. To capture time-dependence, the GPUDDPM allows clusters to `lose assigned observations' in a deletion step, which modifies the sizes $m$ over time. Instead of incorporating cluster size variables $m$ directly, we can formulate an equivalent model which makes use of deletion variables $d$, which represent the time when observations are lost from a cluster . At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables. In particular, we introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed. With these variables, we can now express the $m_{k}$ value at time $t$ as
\begin{equation}
\label{compute_clust_size}
m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
\end{equation}
where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise. Additionally, for an observation $x_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
\begin{equation}
\label{del_rho_form}
l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
\end{equation}
The MCMC sampler described in this section operates in this alternate representation, which we can write, for each time step $t \in \{1, \ldots, T\}$ and clusters $k \in \{ 1, \ldots, K_{t} \} $ at time $t$, as
% \begin{align}
% \begin{split}
% 	 \bold{x}_{i} | c_{i}, \theta_{c_{i},t} &\sim \mathcal{N}(\boldsymbol{\mu}_{c_{i},t}, \Sigma_{c_{i},t})  \mathcal{M}n(\bold{p}_{c_{i},t}) \\
% 	c_{i} | c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha  &\sim  \text{CRP}(c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha) \\
% 	d_{i} | \rho  &\sim \text{Geo}(\rho) + i + 1\\
% 	\theta_{k, t} | \theta_{k, t-1}, \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0}  &\sim 
% \begin{cases}
% 	\text{TK} \hspace{12pt} \text{if } k \in \{ 1, \ldots, K_{t-1}  \} \\
% 	\mathbb{G}_{0}(\boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0})  \hspace{15pt} \text{if } k = K_{t-1} + 1
% \end{cases}
% \end{split}
% \end{align}
% where TK stands for the transition kernel, given in \eqref{transker1} and \eqref{transker2}, and the base distribution $\mathbb{G}_{0}$ is given in \eqref{basedistro}.
\begin{align}
\begin{split}
d_{i,t} | \rho  &\sim \text{Geo}(\rho) + t + 1 \\
\theta_{k, t} | \theta_{k, t-1}  &\sim
\begin{cases}
P(\theta_{k,t} | \theta_{k,t-1}) \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
\mathbb{G}_{0} \hspace{2mm} \text{if} \hspace{2mm} k = K_{t} + 1
\end{cases} \\
c_{i,t} | \bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha  &\sim  \text{CRP}(\bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha) \\
\bold{x}_{i,t} | c_{i,t}, \theta_{c_{i,t},t}  &\sim  F(\theta_{c_{i,t}, t})
\end{split}
\end{align}
where $P(\theta_{k,t} | \theta_{k,t-1})$ is defined in \eqref{transker1} and \eqref{transker2}, $\mathbb{G}_{0}$ is defined in \eqref{basedistro}, CRP is defined in \eqref{CRP}, and $F(\theta_{c_{i,t}, t})$ is defined in \eqref{likelihood}.

This formulation of the GPUDDPM is also used in \cite{gasthaus_thesis} and \cite{caron_2007}. The associated graphical model for this formulation is given in Figure~\ref{fig:gpuddpm_gm_2}.
\begin{figure}[h]
        \center{\includegraphics[width=100mm]{../img/gpuddp_gm_3.pdf}}
        \caption{Graphical model of the deletion variable and auxiliary variable transition kernel representation of the GPUDDPM}
        \label{fig:gpuddpm_gm_2}
\end{figure}

@@

\subsubsection{Sampling Cluster Parameters $\theta_{k, t}$}
The conjugacy present in chosen distributions and auxiliary variables in the transition kernel allow for posterior samples of the cluster parameters to be easily gained, and gives us
% \begin{eqnarray}
% P(\theta_{k,t} | state) && = P(\bold{x}_{i} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
%  && = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
% \end{eqnarray}
\begin{align}
\begin{split}
P(\theta_{k,t} | state) & = P(\bold{x}_{i} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
 & = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{k,t}, \Sigma_{k,t} | \boldsymbol{\mu}_{N}, k_{N}, v_{N}, \Lambda_{N})  \mathcal{D}ir( \bold{p}_{k,t} | \bold{q}_{N})
\end{split}
\end{align}
Where the parameters in the above distribution are given when the data at time $t$, auxiliary variables $z_{k,t-1,1:M}$ at time $t-1$, and auxiliary variables $z_{k,t,1:M}$ at time $t$, are taken to be the observations in the following Bayesian updates
\begin{eqnarray}
\kappa_{N} &=& \kappa_{0} + N \\
\nu_{N} &=& \nu_{0} + N \\
\boldsymbol{\mu}_{N} &=& \frac{\kappa_{0}}{\kappa_{0}+N} \boldsymbol{\mu}_{0}  +  \frac{N}{\kappa_{0}+N} \overline{\bold{x}}^{s}\\
\Lambda_{N} &=& \Lambda_{0} + S_{\bold{x}^{s}}\\
\bold{q}_{N} &=& \bold{q}_{0} + \sum_{i=1}^{N} \bold{x}_{i}^{c}
\end{eqnarray}
where $N$ is the number of observations, $\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{x}^{s}$ and $\bold{x}^{c}$ respectively denote the spatial and color features of the observations, and $\overline{\bold{x}}$ and $S_{\bold{x}}$ respectively denote the sample mean and sample covariance for a set of observations $\bold{x}$, defined in \eqref{samplemean} and \eqref{samplecov}.
% \begin{eqnarray}
% \overline{\bold{x}}  &=&  \left( \sum_{i=1}^{N} \bold{x}_{i} \right) / N\\
% S_{\bold{x}}  &=&  \sum_{i=1}^{N} (\bold{x}_{i} - \overline{\bold{x}}) (\bold{x}_{i} - \overline{\bold{x}})^{T}
% \end{eqnarray}

@@

This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference, which provides another method to carry out unsupervised detection and tracking. The SMC inference algorithm is given in Algorithm~\ref{alg:SMC}. In this algorithm, at each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of sample of the assignment for each observation $\bold{x}_{1,t}, \ldots, \bold{x}_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$. A set of particles is sampled at each time step from relevant importance distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a subset of the particles are resampled and kept. \willie{Gibbs sampling is used to generate the samples associated with each particle at each time step.}

@@

\begin{algorithm}
\caption{Sequential Monte Carlo Inference for the GPUDDPM}
\label{alg:SMC}
\begin{algorithmic}[1]
\FOR{$l = 1 : L$}
\STATE $w_{0}^{(l)} \leftarrow 1/L$ \hfill $\triangleright$ initialize weights
\ENDFOR
\FOR{$t = 1 : T$}
\FOR{$l = 1 : L$}
\STATE $K_{t}^{(l)}  \leftarrow K_{t-1}^{(l)}$
\STATE $m_{1:K_{t}^{(l)}, t}^{(l)} \leftarrow m_{1:K_{t-1}^{(l)}, t-1}^{(l)}$
\FOR{$s = 1 : S$}
\FOR{$i = 1 : N_{t}$}
\IF{$s=1$}
\STATE Sample $c_{i,t}^{(l)} \sim P(c_{i,t}^{(l)} | m_{1:K_{t-1}^{(l)}, t-1}^{(l)}, \theta_{1:K_{t-1}^{(l)}, t-1}^{(l)} )$
\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} + 1$
\ELSE
\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} - 1$
\STATE Sample $c_{i,t}^{(l)} \sim P(c_{i,t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)}, \theta_{1:K_{t}^{(l)}, t}^{(l)} )$  \hfill $\triangleright$ eq.~\eqref{smc_assig_eqn}
\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} + 1$
\ENDIF
\IF{$c_{i,t}^{(l)} = K_{t}^{(l)} + 1$}
\STATE Sample $\theta_{c_{i,t}^{(l)}, t}^{(l)} \sim q_{1}(\bold{x}_{i,t})$ \hfill $\triangleright$ eq.~\eqref{smc_q1_eqn}
\STATE $K_{t}^{(l)} \leftarrow K_{t}^{(l)} + 1$
\STATE $m_{K_{t}^{(l)},t}^{(l)} \leftarrow 1$
\ENDIF
\ENDFOR
\FOR{$k = 1 : K_{t}^{(l)}$}
\IF{$k > K_{t-1}^{(l)}$}
\STATE Sample
\ELSIF{$k \leq K_{t-1}^{(l)}$ and $ \#\{ \bold{x}_{1:N_{t}, t} = k \} > 0$}
\STATE Sample
\ELSIF{$m_{k,t}^{(l)} > 0$}
\STATE Sample
\ENDIF
\IF{$s=S$}
\STATE Sample
\ENDIF
\ENDFOR
\ENDFOR
\STATE Calculate weight of particle
\ENDFOR
\STATE Resample (only keep some of the $L$ particles).
\ENDFOR
\end{algorithmic}
\end{algorithm}
